Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-06-29 manuscript/manuscript.Rmd
updates figures in ms with parameter vary runs

@@

##  Merge IPM and QBM mean parameter results
bothmodel_outs <- rbind(ipm_meansims, qbm_meansims)
meanparam_agg <- ddply(bothmodel_outs, .(species, climsim, model), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.9)
meanplot <- ggplot(meanparam_agg)+
  geom_bar(aes(x=climsim, y=meandiff, fill=model),
             position=dgd, stat="identity")+
  theme_bw()+
  scale_fill_manual(values=c("black", "grey45"), name="Model")+
  facet_wrap("species", scales="free", nrow=1)+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Mean proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

##  Merge IPM and QBM mean parameter results
bothmodel_outs <- rbind(ipm_meansims, qbm_meansims)
meanparam_agg <- ddply(bothmodel_outs, .(species, climsim, model), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.6)
uncertainplot <- ggplot(meanparam_agg)+
  geom_errorbar(aes(x=climsim, ymax=hidiff, ymin=lodiff, group=model, linetype=model),
                position=dgd, width=0.25)+
  theme_bw()+
  scale_linetype_discrete(name="Model")+
  facet_wrap("species", nrow=1, scales="free")+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

>>>>

An important caveat to the above conclusion is that our comparison of the IPM and the QBM

@@

####
##  Mean parameter runs
####
# meanparam_ipm <- data.frame(cover=NA, species=NA, 
#                              climsim=NA, model=NA)
# ##  IPM 
# path <- "./ipm/simulations/results/climchange_meanparams/"
# ipmfiles_mean <- list.files(path)
# ipmfiles_mean <- ipmfiles_mean[grep("cover", ipmfiles_mean)]
# for(file in ipmfiles_mean){
#   tmp <- read.csv(paste(path,file,sep=""))
#   tmp$model <- "IPM"
#   meanparam_ipm <- rbind(meanparam_ipm, tmp[,c("cover", "species", "climsim", "model")])
# }
# meanparam_ipm <- meanparam_ipm[2:nrow(meanparam_ipm),]
# 
# newd <- data.frame(cover=NA, species=NA, 
#                    climsim=NA, model=NA, obscover=NA)
# for(dospp in unique(meanparam_ipm$species)){
#   tmp <- subset(meanparam_ipm, species==dospp)
#   obscover <- subset(tmp, climsim=="obs")["cover"]
#   tmp2 <- subset(tmp, climsim!="obs")
#   tmp2$obscover <- rep(obscover$cover, times = length(unique(tmp2$climsim)))
#   newd <- rbind(newd, tmp2)
# }
# ipm_meansims <- newd[2:nrow(newd),]
# ipm_meansims$propdiff <- with(ipm_meansims, log(cover*100)-log(obscover*100))
# 
# ##  QBM
# meanparam_qbm <- data.frame(cover=NA, species=NA, 
#                              climsim=NA, model=NA)
# path <- "./quadBM/simulations/results/climatechange_meanparams/"
# qbmfiles_mean <- list.files(path)
# for(file in qbmfiles_mean){
#   tmp <- readRDS(paste(path,file,sep=""))
#   tmp$model <- "QBM"
#   meanparam_qbm <- rbind(meanparam_qbm, tmp[,c("cover", "species", "climsim", "model")])
# }
# meanparam_qbm <- meanparam_qbm[2:nrow(meanparam_qbm),]
# 
# newd <- data.frame(cover=NA, species=NA, 
#                    climsim=NA, model=NA, obscover=NA)
# for(dospp in unique(meanparam_qbm$species)){
#   tmp <- subset(meanparam_qbm, species==dospp)
#   obscover <- subset(tmp, climsim=="obs")["cover"]
#   tmp2 <- subset(tmp, climsim!="obs")
#   tmp2$obscover <- rep(obscover$cover, times = length(unique(tmp2$climsim)))
#   newd <- rbind(newd, tmp2)
# }
# qbm_meansims <- newd[2:nrow(newd),]
# qbm_meansims$propdiff <- with(qbm_meansims, log(cover*100)-log(obscover*100))
# 
# ##  Merge IPM and QBM mean parameter results
# bothmodel_outs <- rbind(ipm_meansims, qbm_meansims)
# meanparam_agg <- ddply(bothmodel_outs, .(species, climsim, model), summarise,
#                        meandiff = mean(propdiff),
#                        hidiff = quantile(propdiff, probs = 0.95),
#                        lodiff = quantile(propdiff, probs = 0.05))
# dgd = position_dodge(width = 0.9)
# meanplot <- ggplot(meanparam_agg)+
#   geom_bar(aes(x=climsim, y=meandiff, fill=model),
#              position=dgd, stat="identity")+
#   theme_bw()+
#   scale_fill_manual(values=c("black", "grey45"), name="Model")+
#   facet_wrap("species", scales="free", nrow=1)+
#   scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
#   ylab("Mean proportional difference")+
#   xlab("Climate simulation")+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
#

@@

##  Merge IPM and QBM mean parameter results
bothmodel_outs <- rbind(ipm_meansims, qbm_meansims)

meanparam_agg <- ddply(bothmodel_outs, .(species, climsim, model), summarise,
                       meandiff = mean(propdiff),
                       hidiff = quantile(propdiff, probs = 0.95),
                       lodiff = quantile(propdiff, probs = 0.05))
dgd = position_dodge(width = 0.6)
uncertainplot <- ggplot(meanparam_agg)+
  geom_errorbar(aes(x=climsim, ymax=hidiff, ymin=lodiff, group=model, linetype=model),
                position=dgd, width=0.25)+
  theme_bw()+
  scale_linetype_discrete(name="Model")+
  facet_wrap("species", nrow=1, scales="free")+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

dgd = position_dodge(width = 0.9)
meanplot <- ggplot(meanparam_agg)+
  geom_bar(aes(x=climsim, y=meandiff, fill=model),
             position=dgd, stat="identity")+
  theme_bw()+
  scale_fill_manual(values=c("black", "grey45"), name="Model")+
  facet_wrap("species", scales="free", nrow=1)+
  scale_x_discrete(labels=c("+ppt", "+temp", "+ppt&temp"))+
  ylab("Mean proportional difference")+
  xlab("Climate simulation")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r figure_5, dependson="plot-options", fig.cap="Correlations between QBM and IPM estimates of climate effects. The QBM does not have multiple vital rates, so its values are repeated across panels within each species.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

library(ggmcmc)
library(gridExtra)
library(ggplot2)

allfiles <- list.files("../analysis/speciesData/")
removals <- grep("csv", allfiles)
species_list <- allfiles[-removals]

clim_covs <- c("pptLag", "ppt1", "ppt2", "TmeanSpr1", "TmeanSpr2",
               "ppt1xTmeanSpr1", "ppt2xTmeanSpr2",
               "sizeXpptLag", "sizeXppt1", "sizeXppt2",
               "sizeXTmeanSpr1", "sizeXTmeanSpr2")
clim_mains <- c("pptLag", "ppt1", "ppt2", "TmeanSpr1", "TmeanSpr2",
                "ppt1xTmeanSpr1", "ppt2xTmeanSpr2")

####
##  Get mean growth climate effects
####
growth_parms <- data.frame(species=NA, model=NA, parameter=NA, value=NA)
for(spp in species_list){
  tmp_data <- readRDS(paste("../analysis/ipm/vitalRateRegs/growth/growth_stanmcmc_", spp, ".RDS", sep=""))
  tmp_clim <- tmp_data[grep("b2", tmp_data[,"Parameter"]), ]
  tmp_clim[,"Parameter"] <- rep(clim_covs, each=3000)
  tmp_agg <- ddply(tmp_clim, .(Parameter), summarise,
                   value = median(value))
  tmp_agg2 <- tmp_agg[which(tmp_agg$Parameter %in% clim_mains), ]
  tmp_agg2$species <- spp
  tmp_agg2$model <- "ipm"
  colnames(tmp_agg2) <- tolower(colnames(tmp_agg2))
  growth_parms <- rbind(growth_parms,
                        tmp_agg2[,c("species", "model", "parameter", "value")])
}

####
##  Get mean survival climate effects
####
surv_parms <- data.frame(species=NA, model=NA, parameter=NA, value=NA)
params <- character(length = 12)
for(i in 1:12){
  params[i] <- paste("b2[",i,"]", sep="")
}
for(spp in species_list){
  tmp_data <- readRDS(paste("../analysis/ipm/vitalRateRegs/survival/survival_stanmcmc_", spp, ".RDS", sep=""))
  keeps <- which(tmp_data$Parameter %in% params)
  tmp_clim <- tmp_data[keeps, ]
  tmp_clim[,"Parameter"] <- rep(clim_covs, each=3000)
  tmp_agg <- ddply(tmp_clim, .(Parameter), summarise,
                   value = median(value))
  tmp_agg2 <- tmp_agg[which(tmp_agg$Parameter %in% clim_mains), ]
  tmp_agg2$species <- spp
  tmp_agg2$model <- "ipm"
  colnames(tmp_agg2) <- tolower(colnames(tmp_agg2))
  surv_parms <- rbind(surv_parms,
                        tmp_agg2[,c("species", "model", "parameter", "value")])
}

####
##  Get mean recruitment climate effects
####
rec_parms <- data.frame(species=NA, model=NA, parameter=NA, value=NA)
params <- character(length = 12)
for(i in 1:12){
  params[i] <- paste("b2[",i,"]", sep="")
}
for(spp in species_list){
  tmp_data <- readRDS(paste("../analysis/ipm/vitalRateRegs/recruitment/recruitment_stanmcmc_", spp, ".RDS", sep=""))
  keeps <- which(tmp_data$Parameter %in% params)
  tmp_clim <- tmp_data[keeps, ]
  tmp_clim[,"Parameter"] <- rep(clim_mains, each=3000)
  tmp_agg <- ddply(tmp_clim, .(Parameter), summarise,
                   value = median(value))
  tmp_agg2 <- tmp_agg[which(tmp_agg$Parameter %in% clim_mains), ]
  tmp_agg2$species <- spp
  tmp_agg2$model <- "ipm"
  colnames(tmp_agg2) <- tolower(colnames(tmp_agg2))
  rec_parms <- rbind(rec_parms,
                      tmp_agg2[,c("species", "model", "parameter", "value")])
}

####
##  Get QBM growth climate estimates
####
clim_covs2 <- c("pptLag", "ppt1", "ppt2", "TmeanSpr1", "TmeanSpr2",
               "ppt1xTmeanSpr1", "ppt2xTmeanSpr2",
               "coverXpptLag", "coverXppt1", "coverXppt2",
               "coverXTmeanSpr1", "coverXTmeanSpr2")

qbm_parms <- data.frame(species=NA, model=NA, parameter=NA, qbmvalue=NA)
params <- character(length = 12)
for(i in 1:12){
  params[i] <- paste("b2[",i,"]", sep="")
}
for(spp in species_list){
  tmp_data <- readRDS(paste("../analysis/quadBM/vitalRateRegressions/truncNormModel/popgrowth_stanmcmc_", spp, ".RDS", sep=""))
  keeps <- which(tmp_data$Parameter %in% params)
  tmp_clim <- tmp_data[keeps, ]
  tmp_clim[,"Parameter"] <- rep(clim_covs2, each=3000)
  tmp_agg <- ddply(tmp_clim, .(Parameter), summarise,
                   qbmvalue = median(value))
  tmp_agg2 <- tmp_agg[which(tmp_agg$Parameter %in% clim_mains), ]
  tmp_agg2$species <- spp
  tmp_agg2$model <- "qbm"
  colnames(tmp_agg2) <- tolower(colnames(tmp_agg2))
  qbm_parms <- rbind(qbm_parms,
                     tmp_agg2[,c("species", "model", "parameter", "qbmvalue")])
}

####
##  Format data, combine, and plot
####
growth_parms <- growth_parms[2:nrow(growth_parms), ]
growth_parms$vitalrate <- "growth"
surv_parms <- surv_parms[2:nrow(surv_parms), ]
surv_parms$vitalrate <- "surv"
rec_parms <- rec_parms[2:nrow(rec_parms), ]
rec_parms$vitalrate <- "rec"
ipm_parms <- rbind(growth_parms, surv_parms, rec_parms)

qbm_parms <- qbm_parms[2:nrow(qbm_parms), ]
qbm_parms$vitalrate <- "growth"
vitals <- c("surv", "rec")
qbmalone <- qbm_parms
for(dov in vitals){
  tmp <- qbmalone
  tmp$vitalrate <- dov
  qbm_parms <- rbind(qbm_parms, tmp)
}

ipm_parms <- ipm_parms[,c("species","parameter","value","vitalrate")]
qbm_parms <- qbm_parms[,c("species","parameter","qbmvalue","vitalrate")]
all_ests <- merge(ipm_parms, qbm_parms, by = c("species", "parameter", "vitalrate"))

# Calculate correlation for each group
cors <- ddply(all_ests, c("species", "vitalrate"), summarise, 
              cor = round(cor(value, qbmvalue), 2))

ggplot(all_ests, aes(x=qbmvalue, y=value))+
  geom_point(size=3)+
  geom_smooth(method="lm", color="black", fill="grey")+
  facet_grid(species~vitalrate)+
  geom_text(data=cors, aes(label=paste("r = ", cor, sep="")), x=-1, y=1, size=4)+
  xlab("QBM Estimate")+
  ylab("IPM Estimate")+
  theme_bw()

### 2015-06-29 manuscript/manuscript.tex
updates figures in ms with parameter vary runs

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE).}
\end{figure}

@@

>>>>

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Time series of average percent cover over all quadrats for our
four focal species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa
comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda}
(POSE). Light grey lines show trajectories of individual quadrats. Note
the different y-axis scales across panels.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_5.pdf}
\caption{Correlations between QBM and IPM estimates of climate effects.
The QBM does not have multiple vital rates, so its values are repeated
across panels within each species.}
\end{figure}
